Dynamical Properties of Molecular Cloud Complexes in Galaxies at the Epoch of Reionization

### 2019-03-20 WRITEUP/ms.tex
AF changes

The paper is structured as follows\footnote{Throughout this paper, we adopt a concordance cosmology, with total matter, vacuum and baryonic densities in units of the critical density $\Omega_{\Lambda}$\eq0.692, $\Omega_m$\eq0.308, $\Omega_b$\eq0.0481, Hubble constant $H_0$\eq100\,$h$\,km s\pmOne\,Mpc\pmOne with $h$\eq0.678, spectral index $n$\eq0.967 and $\sigma_8$\eq0.826 \citep{Planck14a}.}. We start by providing some physical background in \Sec{Back}. In \Sec{sim}, we describe the setup of our simulation and properties of our main galaxy (\flower). In \Sec{eqn}, we describe the method used to identify molecular gas complexes and present the formalism within which we interpret the results. In \Sec{results}, we present the results and the scaling relations for the MCs. We then interpret the results and discuss the implications of our findings in \Sec{diss}, and give our conclusions in \Sec{conclusion}.
%
\section{Physical Background}\label{sec:Back}
We begin by introducing some basic material concerning known physical relations and stability analysis of MCs that will become useful in the following.
\subsection{Larson relations}\label{sec:PVE}
\citep{Larson81a} discussed a number of relations among Galactic MC properties, namely the linewidth-size, density-size, and mass-size relations. Larson relations are routinely used for comparing properties of molecular structures in different galactic environments. They also represent a useful framework to analyze our results as they arise from the interplay between gravity and turbulence encoded in the virial theorem.

In general, the virial theorem for a gas distribution of gas can be written as
\begin{equation}\label{eqn:virial_th_general}
\frac{1}{2}\ddot{\mathcal{I}} = 2(\mathcal{T} - \mathcal{T}_{\rm ext}) + \mathcal{W},
\end{equation}
where $\ddot{\mathcal{I}}$ is the second time derivative of the Lagrangian moment of inertia, $\mathcal{T}$ is the internal energy of the gas, $\mathcal{T}_{\rm ext}$ is external pressure support, and $\mathcal{W}$ is the gravitational energy.
%
Let us specialize to the case of a spherical self-gravitating molecular cloud of mass $M_{\rm cl}$, radius $R$, and r.m.s. velocity dispersion $\sigma$, accounting for both thermal and turbulent contributions. Neglecting magnetic field and defining $P_{\rm ext}$ the external pressure, \Eq{virial_th_general} can be written as
\begin{equation}
\frac{1}{2}{\ddot{\mathcal{I}}} = 3 M_{\rm cl} \sigma^2 - 4\pi P_{\rm ext} R^3 - \Gamma\frac{GM_{\rm cl}^2}{R}\,,
\label{eqn:virial}
\end{equation}
where $\Gamma$ is a geometrical factor that is equal to 3/5 for a uniform sphere; in \Eq{virial} the terms on the r.h.s. represent the kinetic, external, and gravity pressure terms. 
%
For the case of simple virial equilibrium (SVE; i.e., equilibrium state without external pressure), \Eq{virial} becomes:
\begin{equation}
\alpha_{\rm vir} \equiv \frac{3\sigma^2}{\Gamma (GM_{\rm cl}/R)} = \frac{5\sigma^2R}{GM_{\rm cl}}\,,
\label{eqn:alpha}
\end{equation}
which defined as the virial parameter $\alpha_{\rm vir}$.

Motivated by the Larson's linewidth-size relation \citep{Larson81a} and the work by \citet{Heyer09a}, we recast Equation~\ref{eqn:alpha} into 
\begin{equation}
\sigma^2/R = \pi G \Sigma/5\,,
\label{eqn:SVE}
\end{equation}
for gas in virial equilibrium, i.e., by setting $\alpha_{\rm vir}$\eq1; the surface density $\Sigma$\eq$M/\pi R^2$. On the other hand, for pressure-bounded virial equilibrium (PVE; i.e. equilibrium state with external pressure), Equation~\ref{eqn:virial} becomes:
\begin{equation}
P_{\rm ext} = \frac{3\sigma^2M_{\rm cl}}{4\pi R^3} - \Gamma\frac{G M_{\rm cl}^2}{4\pi R^4}\,,
\label{eqn:pext}
\end{equation}
which quantifies the external pressure needed to confine the gas inside a given volume. Substituting again $\Sigma$  into Equation~\ref{eqn:pext}, we find 
\begin{equation}
\frac{\sigma^2}{R} = \frac{1}{3}\left(\frac{4P_{\rm ext}}{\Sigma} + \Gamma G \Sigma \pi \right)\,,
\label{eqn:v0}
\end{equation}
keeping $\sigma^2/R$ on the LHS.

Based on Equations~\ref{eqn:SVE} and \ref{eqn:pext}, a one-to-one mapping between $\sigma^2/R$ and $\Sigma$ is therefore expected for a virialized cloud, since $\sigma^2/R\propto\Sigma$. Equation~\ref{eqn:v0} then yields the loci along which external pressures $P_{\rm ext}$ are needed in order for some given MCs to have linewidths $\sigma$ for a given set of surface densities. In other words, a cloud with $\sigma$ and $\Sigma$ offset from $\sigma^2/R\mapsto\Sigma$ is in equilibrium with external pressure $P_{\rm ext}$; therefore $\sigma^2/R$ is not constant with $\Sigma$ but varies depending on $P_{\rm ext}$ (cf. expectation for SVE, where $\sigma\propto R^{0.5}$; see e.g., \citealt{Heyer09a, Hughes10a, Hughes13b, Meidt13a}).

Summarizing, the virial parameter can be used to quantify the stability/boundedness of an MC. Accounting for the external pressure, a virial parameter of $\alpha_{\rm vir}\lesssim2$ would be unstable \citep{bertoldi:1992}.
%
Such criterion is often used in \obs \citep[see e.g., ][]{Kauffmann17b}; in the following, to facilitate comparison with \obs, we derive $\alpha_{\rm vir}$ using the gas component only and ignore contributions due to the stellar component.

@@

The onset of gravitational instability is considered to be tightly connected to \SF \citep[e.g.,][]{Kennicutt89a, Wang94a, Li05b, Li06a}. From virial theorem, Jeans instability sets the scales at which gravitational potential overcomes the internal energy. Namely, gravitational instabilities dominates at scales $L > L_J$, where $L_J$ is the Jeans length. However, centrifugal force resulting from differential rotation in disk galaxies can inhibit collapse even on these scales. As a result, a disk is thought to be unstable on scales between $L_J < L < L_{\rm rot}$, where $L_{\rm rot}$ is set by the surface density of the disk and the angular velocity of the rotation.

In theoretical and observational works, gravitational instability has been widely discussed in the context of the Toomre $Q$ stability criterion \citep{Toomre64a, Goldreich65b} or its observable proxy  $v_{\rm circ}/\sigma$, i.e., the circular velocity to dispersion ratio \citep[e.g.,][]{GarciaBurillo03a, Genzel11a, Kassin12a, Leung19a}.
% https://ned.ipac.caltech.edu/level5/March15/Glazebrook/Glazebrook5.html
%
% from linear analysis
For {\it axisymmetric} mode, the dispersion relation for the growth of density perturbation in a rotating turbulent disk of finite thickness $h$ is described by
\begin{equation}
\omega^2 = \kappa^2 - \frac{2\pi G \Sigma |k|}{1 + |k| h} + \sigma^2 k^2\,,
\label{eqn:3Ddisp}
\end{equation}
where $k$ is the wavenumber and $\kappa$ is the epicyclic frequency, defined as:
\begin{equation}
\kappa^2\equiv\frac{2\Omega}{R}\frac{d}{dR}\left(R^2\Omega\right)\,.
\label{eqn:kappa}
\end{equation}
\citep{Romeo92a}.
The first term on the RHS of Equation~\ref{eqn:3Ddisp} describes rotation support, whereas the second term describes self-gravity, and the third term describes internal pressure support. The Toomre $Q$ parameter for the gas component can be defined as \citep{Toomre64a}:
\begin{equation}
Q_{\rm gas}\equiv\frac{\sigma_{\rm gas}\kappa}{\pi G \Sigma_{\rm gas}}\,.
\label{eqn:Q}
\end{equation}
In the thin disk approximation ($kh\ll1$), the solution of \Eq{3Ddisp} is unstable for scales $k$ such that $Q < Q_{\rm crit}\simeq1$ (or equivalently $\omega^2 < 0$). Physically, this situation occurs when local gravitational instability
overcomes support from pressure and rotation, allowing fragmentation and collapse to happen. Similarly, for a collisionless fluid --- as an ensemble of stars --- the Toomre $Q$ parameter is defined by \citep{Toomre64a}:
\begin{equation}
Q_{\star} \equiv\frac{\sigma_{\star}\kappa}{3.36 G \Sigma_{\star}}\,.
\end{equation}

@@

\subsection{Properties of the Clouds}

>>>>

The paper is structured as follows\footnote{Throughout this paper, we adopt a concordance cosmology, with total matter, vacuum and baryonic densities in units of the critical density $\Omega_{\Lambda}$\eq0.692, $\Omega_m$\eq0.308, $\Omega_b$\eq0.0481, Hubble constant $H_0$\eq100\,$h$\,km s\pmOne\,Mpc\pmOne with $h$\eq0.678, spectral index $n$\eq0.967 and $\sigma_8$\eq0.826 \citep{Planck14a}.}. We start by providing some physical background in \Sec{Back}. In \Sec{sim}, we describe the setup of our simulation and properties of our main galaxy (\flower). In \Sec{eqn}, we describe the method used to identify molecular gas complexes and present the formalism within which we interpret the results. In \Sec{results}, we present the results and the scaling relations for the MCs. We then interpret the results and discuss the implications of our findings in \Sec{diss}, and give our conclusions in \Sec{conclusion}.
%
\section{Physical Background}\label{sec:Back}
We begin by introducing some basic material concerning notable physical ("Larson") relations characterizing MCs, and the gravitational instability of galactic disks, which might be driving MC formation. Such notions will be used in the subsequent analysis of our simulations. 
\subsection{Larson relations and virial parameter}\label{sec:PVE}
\citep{Larson81a} discussed a number of relations among Galactic MC properties, namely the linewidth-size, density-size, and mass-size relations. Larson relations are routinely used for comparing properties of molecular structures in different galactic environments. They also represent a useful framework to analyze our results as they arise from the interplay between gravity and turbulence encoded in the virial theorem.

In general, the virial theorem for a gas distribution of gas can be written as
\begin{equation}\label{eqn:virial_th_general}
\frac{1}{2}\ddot{\mathcal{I}} = 2(\mathcal{T} - \mathcal{T}_{\rm ext}) + \mathcal{W},
\end{equation}
where $\ddot{\mathcal{I}}$ is the second time derivative of the Lagrangian moment of inertia, $\mathcal{T}$ is the internal energy of the gas, $\mathcal{T}_{\rm ext}$ is external pressure support, and $\mathcal{W}$ is the gravitational energy.
%
Let us specialize to the case of a spherical self-gravitating molecular cloud of mass $M_{\rm cl}$, radius $R$, and r.m.s. velocity dispersion $\sigma$, accounting for both thermal and turbulent contributions. Neglecting magnetic field and defining $P_{\rm ext}$ the external pressure, \Eq{virial_th_general} can be written as
\begin{equation}
\frac{1}{2}{\ddot{\mathcal{I}}} = 3 M_{\rm cl} \sigma^2 - 4\pi P_{\rm ext} R^3 - \Gamma\frac{GM_{\rm cl}^2}{R}\,,
\label{eqn:virial}
\end{equation}
where $\Gamma$ is a geometrical factor that is equal to 3/5 for a uniform sphere; in \Eq{virial} the terms on the r.h.s. represent the kinetic, external, and gravity pressure terms.

Motivated by Larson's linewidth-size relation \citep{Larson81a} and the work by \citet{Heyer09a}, we assume equilibrium (i.e. ${\ddot{\mathcal{I}}}=0$), define the cloud surface density as
$\Sigma$\eq$M/\pi R^2$, and rewrite previous equation as 
\begin{equation}
\frac{\sigma^2}{R} = \frac{1}{3}\left(\frac{4P_{\rm ext}}{\Sigma} + \frac{3}{5} \pi G \Sigma \right)\,,
\label{eqn:v0}
\end{equation}
which further reduces to 
\begin{equation}
\frac{\sigma^2}{R} = \frac{\pi}{5} G \Sigma\,,
\label{eqn:SVE}
\end{equation}
if the external pressure $P_{\rm ext}=0$. For this case (often referred to as simple virial equilibrium) from the balance between kinetic and gravity terms we can define the virial parameter as 
\begin{equation}
\alpha_{\rm vir} \equiv  \frac{5\sigma^2R}{GM_{\rm cl}} = \frac{5\sigma^2}{\pi G \Sigma R}\,,
\label{eqn:alpha}
\end{equation}

Based on Equation~\ref{eqn:SVE}, a one-to-one mapping between $\sigma^2/R$ and $\Sigma$ is therefore expected for a virialized cloud, since $\sigma^2/R\propto\Sigma$. Deviations from this relation are usually attributed to a significant contribution from external pressure as per Equation~\ref{eqn:v0} (see e.g., \citealt{Heyer09a, Hughes10a, Hughes13b, Meidt13a}).

Summarizing, the virial parameter can be used to quantify the stability/boundedness of an MC. Accounting for the external pressure, a virial parameter of $\alpha_{\rm vir}\lesssim2$ would be unstable \citep{bertoldi:1992}.
%
Such criterion\footnote{In the following, to facilitate comparison with \obs, we derive $\alpha_{\rm vir}$ using the gas component only and ignore contributions due to the stellar component.} is often used in \obs \citep[see e.g., ][]{Kauffmann17b}.

@@

The onset of gravitational instability is considered to be tightly connected to \SF \citep[e.g.,][]{Kennicutt89a, Wang94a, Li05b, Li06a}. From virial theorem, Jeans instability sets the scales at which gravitational potential overcomes the internal energy. Namely, gravitational instabilities dominates at scales $L > L_J$, where $L_J$ is the Jeans length. However, centrifugal forces resulting from differential rotation in disk galaxies can stabilize perturbations even on these scales. As a result, a disk is thought to be unstable on scales between $L_J < L < L_{\rm rot}$, where $L_{\rm rot}$ is set by the surface density of the disk and the angular velocity of the rotation.

In theoretical and observational works, gravitational instability has been widely discussed in the context of the Toomre $Q$ stability criterion \citep{Toomre64a, Goldreich65b} or its observable proxy  $v_{\rm circ}/\sigma$, i.e., the circular velocity to the disk r.m.s. velocity dispersion \AF{This is not exactly the same $\sigma$ as before} ratio \citep[e.g.,][]{GarciaBurillo03a, Genzel11a, Kassin12a, Leung19a}.
% https://ned.ipac.caltech.edu/level5/March15/Glazebrook/Glazebrook5.html
%
% from linear analysis
For axisymmetric modes, the dispersion relation for the growth of density perturbations in a rotating, turbulent disk of finite thickness $h$ is described by
\begin{equation}
\omega^2 = \kappa^2 - \frac{2\pi G \Sigma |k|}{1 + |k| h} + \sigma^2 k^2\,,
\label{eqn:3Ddisp}
\end{equation}
where $k$ is the wavenumber and $\kappa$ is the epicyclic frequency, defined as:
\begin{equation}
\kappa^2\equiv\frac{2\Omega}{R}\frac{d}{dR}\left(R^2\Omega\right)\,.
\label{eqn:kappa}
\end{equation}
\citep{Romeo92a}.
\AF{$R$ is not the same as the MC radius in the proev. Section}
In Equation~\ref{eqn:3Ddisp} the terms on the r.h.s. are related to rotation, self-gravity and internal pressure, respectively. The Toomre $Q$ parameter for the gas component can be defined as \citep{Toomre64a}:
\begin{equation}
Q_{\rm gas}\equiv\frac{\sigma_{\rm gas}\kappa}{\pi G \Sigma_{\rm gas}}\,.
\label{eqn:Q}
\end{equation}
In the thin disk approximation ($kh\ll1$), instability ($\omega^2 < 0$ in  \Eq{3Ddisp}) occurs on scales $k$ such that $Q < Q_{\rm crit}\simeq1$ (or equivalently $\omega^2 < 0$). Physically, this implies that the local gravitational field 
overcomes support from pressure and rotation, allowing fragmentation and collapse to happen. Similarly, for a collisionless fluid --- as an ensemble of stars --- the Toomre $Q$ parameter is defined by \citep{Toomre64a}:
\begin{equation}
Q_{\star} \equiv\frac{\sigma_{\star}\kappa}{3.36 G \Sigma_{\star}}\,.
\end{equation}

@@

\subsection{Molecular Cloud Properties}
